---
title: "Social Network Dispersion Activity"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE) #remove eval = FALSE when using this markdown
```

```{r}
library(jsonlite)
library(dplyr)
library(purrr)
library(igraph)
```

### Goal

As we have learned, measuring network dispersion between two individuals is a decent way of identifying whether a romantic relationship may be present. In this activity, you will implement functions to calculate the dispersion value between two vertices.

If you finish early, there is an additional challenge section where you will finish implementing the recursive dispersion function, and calculate the maximum dispersion to predict the partner of a user.

### Collect Data

To get started, collect your own data from https://lostcircles.com/. Download the data (select JSON(no pics)) after loading your network. Run the following chunk to get the nodes and links data frames. Simply fill in your own file path to generate the data frames: 

```{r}
data <- fromJSON("YOUR_FILE_PATH_TO_LOSTCIRCLES_FILE_HERE.json")

links <- data$links %>%
  mutate(source = source + 1, target = target + 1)

nodes <- data$nodes %>% 
  mutate(full_name = name, name = seq(1, nrow(nodes))) %>%
  select(name, full_name)
```

`name` is the attribute used by `igraph` as the index for the vertices.

### Selecting Some Friends

Next you should select two close friends from your Facebook network. In this section we are going to attempt to calculate the dispersion value between two individuals. 

For the best results, the two friends you select should have a high amount of mutual friends who you are also mutual friends with. 

To find the indices of the friends you selected, you can run the following chunk:

```{r}
nodes %>% filter(full_name %in% c("friend_1", "friend_2"))
```

### Core `igraph` Functions

Generate the graph from the links and nodes tables using `igraph`:

```{r}
fbg <- graph_from_data_frame(links, directed = FALSE, vertices = nodes)
```

Get the direct neighbors of a vertex (the list returned includes the vertex itself):

```{r}
ego(fbg, 1, 31)[[1]]
```

Notice that the treshold in this case is 1, meaning we are looking for vertices 1 edge away from vertex 31. If we were interested in all vertices, one or two edges away from vertex 31, the threshold will be 2, and the command to run is `ego(fbg, 2, 31)[[1]]`.

**Task**: Find the direct neighbors of the two friends you selected. Note: only the indices of the vertices that represent your friends should be used.

```{r}
# --- Complete this chunk ---

```

Find the common neighbors of two vertices:

```{r}
intersection(neighbors(fbg, 31), neighbors(fbg, 12))
```

**Task**: Find the common friends of the two friends you selected.

```{r}
# --- Complete this chunk ---
```

### Calculating Dispersion Between Two Vertices

In this section we will calculate the dispersion between the two friends you have found. Your task is to complete these two algorithms. Once you have done that you will be able to find the dispersion value between your selected friends.

**Task**: Implement the distance function

```{r}
# --- Complete this function ---
distance <- function(graph, v1, v2, threshold = 1) {
  # @param graph: the graph generated using igraph
  # @param v1, v2: the vertices to compute the distance for. (int)
  # @param threshold: the distance threshold, 1 by default. (int)
  
  # Use the igraph functions to find the neighbors of v2 with respect to the treshold
  # Check if v1 belongs to the neighbors of v2
  # If v1 does not belong to the neighbors of v2 (distance between v1 and v2 above threshold), return 1
  # Otherwise, return 0
  
  # ...
}
```

**Task**: Implement the dispersion function

```{r}
# --- Complete this function ---
dispersion <- function(graph, v1, v2, threshold = 1, normalized = FALSE) {
  # @param graph: the graph generated using igraph
  # @param v1, v2: the vertices to compute the dispersion for (int)
  # @param threshold: the distance threshold, 1 by default (int)
  # Loop over all the common neighbors of v1 and v2 and return the sum of their distances
  # When normalized, the sum of distances is divided by the number of common neighbors
  # If the vertices have one or no common neighbors, the dispersion should be set to 0
  cn <- intersection(neighbors(graph, v1), neighbors(graph, v2))
  if (length(cn) <= 1) return(0)
  # ...
}
```

**Testing**: make sure you get the correct value for each one of the following tests. The correct test values are shown in the comments above each command.

````{r}
testing_links <- read.csv("https://raw.githubusercontent.com/abelemlih/ci_fb_dispersion/master/testing_graph.csv", header = TRUE, stringsAsFactors = FALSE)
testing_nodes <- data.frame(name = c(letters[1:11], 'u'))
testing_graph <- graph_from_data_frame(testing_links, directed = FALSE, vertices = testing_nodes)
```     

```{r}
# distance between u and h should equal 0 (threshold = 1)
distance(testing_graph, V(testing_graph)['u'], V(testing_graph)['h'], threshold = 1)
```

```{r}
# distance between a and h should equal 1 (threshold = 1)
distance(testing_graph, V(testing_graph)['a'], V(testing_graph)['h'], threshold = 1)
```

```{r}
# distance between c and j should equal 0 (threshold = 2)
distance(testing_graph, V(testing_graph)['c'], V(testing_graph)['j'], threshold = 2)
```

**Task**: Find the dispersion value of the vertices that represent the friends you selected at the start of the activity

```{r}
# --- Complete this chunk ---
```

### Challenge: Calculate max dispersion using recursive dispersion 

In this section you will calculate the maximum dispersion for one of your friends in your network. To do this, fill out the functions below. We have provided a helper function to facilitate dispersion calculation. Computing the recursive dispersion is slower in R, so we recommend being patient especially for very large graphs.

* The number of recursion iterations determines how many recursions our recursive dispersion function will undergo. The paper we read stated that 3 recursions provides a good balance between accuracy and speed.

* The distance threshold determines how many degrees of separation two nodes must be to not be considered neighbors. A higher threshold means you are only looking at nodes where mutual friends are further apart.

**Task** Finish implementing the recursive dispersion function:

```{r}
rdisp <- function(graph, df, v1, v2, threshold = 1) {
  # This function computes the recursive dispersion between two vertices
  # The last column of the data frame df must contain dispersion values from the last recursive iteration
  # @param graph: the graph generated using igraph
  # @param df: the dataframe that contains all dispersion values from previous iterations
  # This dataframe is initialized with one column of iteration values equal to 1
  # @param v1, v2: the vertices to compute the dispersion for
  # @param threshold: the distance threshold, 1 by default
  cn <- intersection(neighbors(graph, v1), neighbors(graph, v2))
  cn_num <- length(cn)
  last_col <- ncol(df)
  result <- 0
  if (cn_num <= 1) return(0)
  
  for(i in seq(1, cn_num - 1)) {
    result <- result + (df[cn[i], last_col])^2
    for(j in seq(i + 1, cn_num)) {
      p <- c(df[cn[i], last_col], df[cn[j], last_col]) %>% prod()
      result <- result + 2*p*distance(graph, cn[i], cn[j], threshold)
    }
  }
  (result + (df[cn[cn_num], last_col])^2)/cn_num
}

# --- Complete this function ---
recursive_dispersion <- function(graph, u, max_iter = 1, threshold = 1) {
  # @param graph: the graph generated using igraph
  # @param u: the vertex to compute the recursive dispersion for
  # @param max_iter: the number of iterations to run
  # @param threshold: the distance threshold, 1 by default
  df <- data.frame(v1 = u, v2 = nodes$name, iter0 = 1)
  for (i in 1:max_iter) {
    # iterate through every vertex in the column v2 of the dataframe df
    # compute the dispersion between that vertex and the main vertex u using the function above
    # add the all the dispersion calculations as a new column in df
    # ...
  }
  df %>% filter(v2 != u)
}
```

**Task** Find the friend with the maximum recursive dispersion:

```{r}
# Uncomment when recursive_dispersion is complete
# dispersions <- recursive_dispersion(fbg, 12, max_iter = 1)

# --- Complete this chunk ---
# Filter the dispersions data frame to find the vertex with highest dispersion value
# Only consider the values from the last column in the data frame, which matches the values from the last recursion iteration
# For instance, if only one recursion iteration was processed, 
# you should filter using the dispersion values from iter1.
```





